---
title: \sffamily{\bfseries{\LARGE My title goes here}} 
author: "Steven J. Pierce"
output:
 pdf_document:
   latex_engine: xelatex
   number_sections: true
   toc: yes
   toc_depth: 3
   includes:
     in_header: "compact-title.tex"
urlcolor: blue
params:
 LogFile: "Example_Script.pdf"
header-includes:
- \usepackage{fancyhdr}
- \usepackage[yyyymmdd,hhmmss]{datetime}
- \usepackage{lastpage}
- \usepackage{fontspec}
- \defaultfontfeatures{Ligatures=TeX}
- \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
- \usepackage{url}
- \usepackage{floatrow}
- \floatplacement{figure}{!ht}
- \floatplacement{table}{!ht}
- \usepackage{placeins}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{dcolumn}
- \usepackage{titling}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \pretitle{\begin{center}\LARGE\bfseries}
- \posttitle{\end{center}}
- \pagestyle{fancy}
- \lhead{Left header text }
- \rhead{\today\ \currenttime}
- \lfoot{\texttt{\small \detokenize{`r params$LogFile`}}} 
- \cfoot{ }
- \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
- \renewcommand{\headrulewidth}{0.4pt}
- \renewcommand{\footrulewidth}{0.4pt}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

# Heading Level 1
Put text here

\FloatBarrier

# Setup
Set global R chunk options (local chunk options will over-ride global options). 
The method for creating a size option that controls font size in code chunks and
their text output is based on an answer to a question posted on 
[stackoverflow.com](https://stackoverflow.com/a/46526740). 

``` {r global-options, cfsize = "footnotesize"}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE, cfsize = "footnotesize")

# Declare location of this script relative to the project root directory.
here::i_am(path = "inst/Example_Script.Rmd")
```

Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(here)         # for here()
library(dplyr)        # for %>%, arrange(), filter(), group_by(), mutate(), etc.
library(tidyr)        # for pivot_longer(), pivot_wider(), etc.
library(rmarkdown)    # for render()
library(knitr)        # for kable()
opts_knit$set(eval.after = 'fig.cap') 
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)   # for kable_styling(), add_header_above(), column_spec(),
                      # collapse_rows(), and landscape()
library(piercer)      # for git_report()
```

\FloatBarrier

# Heading Level 1
Here is is an example of how to get dynamic table and figure number referencing,
plus nice captions and footnotes. We can use the name of the chunk that creates 
the table as follows. Table \ref{tab:example-table} is very simple, and so is 
Figure \ref{fig:example-figure}. 

```{r example-table}
TC <- "Here's A Table Caption."
FN <- "This is the footnote."

# Create som data.
MyData <- data.frame(x = 1:5, y = 2:6, z = 3:7, a = 4:8) 
# Create a table and add a footnote. 
kable(MyData, format = "latex", booktabs = TRUE, caption = TC) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```


```{r example-figure, fig.width=6, fig.height=4, fig.cap=FC}
FC <- "\\label{fig:example-figure} Here's A Figure Caption."
plot(MyData)
```

We can use `\FloatBarrier` commands in front of section headings to force 
floating tables and figures to stay in the section where they were created. 

\FloatBarrier

## Heading Level 2
Put text here. 

\FloatBarrier

# Software Information
We use R Markdown to enhance reproducibility. Knitting the source R Markdown 
script *`r knitr:::current_input()`* generates this PDF file containing 
explanatory text, R code, plus R output (text and graphics).

- We used [RStudio](www.rstudio.org) to work with R and R markdown files. 
- Our software chain looks like this:
  **Rmd file > RStudio > R > rmarkdown > knitr > md file > pandoc > tex file > TinyTeX > PDF file**.
- Source script: *`r knitr:::current_input()`*
- Output file: *`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`*
- We recommend using [TinyTeX](https://yihui.org/tinytex/) to compile LaTeX files 
  into PDF files. However, it should be viable to use 
  [MiKTeX](https://miktex.org) or another LaTeX distribution instead. 
- We used [pandoc](https://pandoc.org) `r pandoc_version()` for this 
  document. 

This document was generated using the following computational environment and 
dependencies: 

``` {r show-citations, cfsize = "scriptsize", eval = TRUE}
# Check and report whether we used TinyTex or MiKTeX. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

The current Git commit details and status are:

```{r git-report, echo = TRUE, eval = TRUE}
git_report()
```
