---
title: \bfseries{Example Script for Rendering PDF Output\LARGE}
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceDir: "scripts/"
  SourceFile: "Example_Render_to_PDF.qmd"
  LogFile: "Example_Render_to_PDF.pdf"
format: 
  pdf:
    output-file: "Example_Render_to_PDF.pdf"
    output-ext: "pdf"
    documntclass: scrartcl
    papersize: letter
    geometry: 
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    colorlinks: true
    template-partials: 
      - title.tex
    include-in-header: 
      - file: compact-title.tex
      - text: | 
          \usepackage{fancyhdr}
          \usepackage[noblocks]{authblk}
          \renewcommand*{\Authsep}{, }
          \renewcommand*{\Authand}{, }
          \renewcommand*{\Authands}{, }
          \renewcommand\Affilfont{\small}
          \usepackage[yyyymmdd,hhmmss]{datetime}
          \usepackage{lastpage}         %% For pageref command
          \usepackage{fontspec}
          \defaultfontfeatures{Ligatures=TeX}
          \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
          \usepackage{url}
          \usepackage{verbatim}         %% For the comment command
          \usepackage{placeins}         %% For FloatBarrier command
          \usepackage{booktabs}         %% Used by kableExtra
          \usepackage{longtable}        %% Used by kableExtra
          \usepackage{array}            %% Used by kableExtra
          \usepackage{multirow}         %% Used by kableExtra
          \usepackage{wrapfig}          %% Used by kableExtra
          \usepackage{colortbl}         %% Used by kableExtra
          \usepackage{pdflscape}        %% Used by kableExtra
          \usepackage{tabu}             %% Used by kableExtra
          \usepackage[normalem]{ulem}   %% Used by kableExtra
          \usepackage{makecell}         %% Used by kableExtra
          \usepackage{xcolor}           %% Used by kableExtra
          \usepackage{dcolumn}
          \usepackage{titling}
          \usepackage{threeparttable}   %% Used by kableExtra
          \usepackage{threeparttablex}  %% Used by kableExtra
          \pretitle{\begin{center}\LARGE\bfseries}
          \posttitle{\end{center}}
          \pagestyle{fancy}
          \lhead{Left header text }
          \rhead{\today\ \currenttime}
          \cfoot{ }
          \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
          \renewcommand{\headrulewidth}{0.4pt}
          \renewcommand{\footrulewidth}{0.4pt}
          \fancypagestyle{plain}{\pagestyle{fancy}}
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true
    cfsize: "scriptsize"
---

\begin{comment}
# Hidden: Notes for User
This script makes use of the LaTeX `verbatim` package's `\comment` command to
build in hidden headings and text that are here for the team preparing the
report but will be omitted from the knitted PDF output file. This makes it
easier to put important comments directly in the file while still creating fully
formatted output that only shows the desired content. So, text inside a comment
block will be omitted from the PDF output.

We can also use code chunks with chunk options such as `#| include: false`, 
`#| echo: false`, `#| message: false`, `#| warning: false` and so on to 
precisely control what R output shows up in the final PDF file. Thus, a script 
file can both serve as process documentation and efficiently generate the final 
version of a report output by applying the reproducible research and literate
programming concepts and tools. 

The `\lfoot` command just below sets the left footer to match the script's 
`LofgFile` parameter value, which we can pass along when rendering the script. 
Ideally it would be in the YAML `include-in-header:` with other pieces, but
I can't get inline code to work if I do it that way. This is the only way I 
could get the filename-based footer to work. 

The LaTeX `placeins` pacakg provides the `\FloatBarrier` command. Using that 
before section headings help keep the figures and tables in their respective 
sections. 
\end{comment}

\lfoot{\texttt{\small \detokenize{`r params$LogFile`}}} 

\FloatBarrier

# Purpose
This script provides a template example you can use to jump-start creating a 
new report that writes PDF output. It illustrates some convenient features of
Quarto.

\FloatBarrier

# Setup
Global R chunk options are defined in the YAML header but local chunk options 
will over-ride global options. We can temporarily disable an individual chunk by 
inserting `#| eval: false` on a line at the top of the chunk. The method for
creating a `cfsize` option that controls font size in code chunks and their text
output is based on an answer to a question posted on
[stackoverflow.com](https://stackoverflow.com/a/46526740).

``` {r}
#| label: global-options
#| message: true

# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", 
         paste0("\n \\", options$cfsize,"\n\n", x, "\n\n \\normalsize"), 
         x)
  })
```

``` {r}
#| label: test-message1

# This message should show in results because of global options in YAML. 
message("test message 1")
```

``` {r}
#| label: test-message2
#| message: false

# This message should not show in results because of chunk-level options. 
message("test message 2")
```

\FloatBarrier

## Declare Path
This next chunk declares the path to this script relative to the project-level 
root directory. If the file is not in the right location under the project root
you'll get a warning message. This helps ensure relative paths are all working 
as expected. 

``` {r}
#| label: declare-path

# Declare path to this script relative to the project root directory.
here::i_am(path = paste0(params$SourceDir, params$SourceFile))
```

\FloatBarrier

## Load Packages
Next we load R packages required for this script to work. 

```{r}
#| label: load-packages

library(devtools)       
library(here)         # for here(), i_am()
library(dplyr)        # for %>%, arrange(), filter(), group_by(), mutate(), etc.
library(tidyr)        # for pivot_longer(), pivot_wider(), etc.
library(rmarkdown)    # for render()
library(knitr)        # for kable()
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(piercer)      # for git_report(), which_latex().
library(quarto)       # for quarto_version()
```

\FloatBarrier

# Heading Level 1
Put text here. 

\FloatBarrier

## Heading Level 2
Put text here. 

\FloatBarrier

### Heading Level 3
Put text here. 

\FloatBarrier

#### Heading Level 4
The `toc-depth: 3` option in YAML means this heading is omitted from the side 
menu. The `number-depth: 3` option in YAML is why this heading lacks a section
number. 

\FloatBarrier

# Examples of Callouts
These are examples of how to use callouts in Quarto documents. 

::: {.callout-note}
Note that there are five types of callouts, including:
`note`, `warning`, `important`, `tip`, and `caution`.
:::

::: {.callout-important}
This is an example of an important callout without a caption.
:::

::: {.callout-warning}
This is an example of an warning callout without a caption.
:::

::: {.callout-tip}
# Tip With Caption
This is an example of a tip callout with a caption.
:::

::: {.callout-caution}
# Caution: Collapse option does not work in PDF output
In PDF output, callouts created with `collapse="true"` just get expanded as if
you had used `collapse="false"`. There is no way to make an interactive, 
expandable callout in this output format.
:::

\FloatBarrier

# Example Table
See @tbl-Test for an example of the table created below. 

```{r}
#| label: tbl-Test
#| tbl-cap: Table Caption Via Chunk Option. 
FN <- "This is the footnote."

# Create some data.
MyData <- data.frame(x = 1:5, y = 2:6, z = 3:7, a = 4:8) 

# Create a table and bold the first row, and add a footnote. 
kable(MyData, format = "latex", booktabs = TRUE) %>% 
  row_spec(row = 1, bold = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = FALSE)
```

::: {.callout-tip}
## Tip: Setting the Table Caption
Use the `tbl-cap` chunk option to set the table caption rather than using the 
`caption = ` argument to `kable()`. The latter causes rendering problems. 
:::

::: {.callout-caution}
## kableExtra in PDF output
It looks like kableExtra footnotes work in PDF output from Quarto documents if 
`threeparttable = FALSE`, but that makes the table wider rather than wrapping 
long footnotes to match the table width. Setting `threeparttable = TRUE` breaks 
rendering. The latter setting works fine in R Markdown documents.
:::

\FloatBarrier

# Example Figure
Below is @fig-example, which is pretty simple. 

```{r}
#| label: fig-example
#| fig-cap: !expr paste("Here's a Computed", 
#|                      "Figure Caption.")
#| fig-width: 6
#| fig-height: 4

plot(MyData)
```

::: {.callout-tip}
## Tip: Setting the Figure Caption
Use the `fig-cap` chunk option to set the figure caption. Here you can see the 
use of `!expr` preceding inline r code that computes a dynamic caption. 
:::

\FloatBarrier

# Example Graphvis Diagram
Here is an example diagram. Notice that it does not get a Figure number by 
default. We can also just use the `DiagrammeR` package to generate a figure 
that gets numbered. 

```{dot}
digraph D {
  A -> {B, C, D} -> {F}
}
```

\FloatBarrier

# Software Information
We used [R](https://www.r-project.org/) as our main computing environment and 
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used 
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto. 

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > tex file > TinyTeX > PDF file**.
- Source file: \texttt{\detokenize{`r params$SourceFile`}} 
- Output file: \texttt{\detokenize{`r params$LogFile`}} 
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.
- [TinyTeX](https://yihui.org/tinytex/) compiles LaTeX files (`*.tex`) into PDF 
  files. It should be viable to use [MiKTeX](https://miktex.org) or another 
  LaTeX distribution instead. 

\FloatBarrier

## Versions
This document was generated using the following computational environment and 
dependencies: 

``` {r}
#| label: show-version
#| cfsize: scriptsize
#| echo: true

# Check and report whether we used TinyTex or other LaTeX software. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

\FloatBarrier

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| cfsize: scriptsize
#| echo: true

git_report()
```
