---
title: "Setup a Research Compendium R Package"
author: "Steven J. Pierce"
date: "`r format(Sys.time(), '%Y-%m-%d, %H:%M:%S %Z')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding:  show
---

# Purpose
This [R Markdown](https://rmarkdown.rstudio.com/) script illustrates the process
of taking a new repository created from the research compendium R package
template located at https://github.com/sjpierce/rcrpkg.template.git and turning
it into a full R package. It is intended to be a script that you use
interactively. Examine each section below, look at what the next chunk does,
edit is desired, and then run it. You will need to manually edit some of the
files created by certain chunks below.

[Quarto](https://quarto.org/) is the successor to 
[R Markdown](https://rmarkdown.rstudio.com/). It offers more features and so I
have converted over to using Quarto. There are examples of both Quarto 
(`*.qmd` files) and R Markdown scripts (`*.Rmd` files) in this repository but I now
prefer and recommend relying only on Quarto. If you intend to do that, you can
delete the (`*.Rmd` files) from your repository. The R Markdown files are only
here for historical examples of the older approach.

# Setup
Set global R chunk options (local chunk options will over-ride global options). 
We can temporarily disable an individual chunk by setting `eval = FALSE`. 

``` {r global-options}
# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, eval = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE)

# Declare path to this script relative to the project root directory.
here::i_am(path = "inst/Setup_as_Package.Rmd")
```

```{r install-packages-1, eval=FALSE}
# Packages required by the functions defined in "R/Utility_Functions.R"
install.packages("assertthat")  
install.packages("git2r")  
install.packages("here")  
install.packages("tinytex") 
```

```{r install-packages-2, eval=FALSE}
# Packages we load below (that are not already in the list above)
install.packages("available") 
install.packages("devtools")
install.packages("fs")
install.packages("here")  
install.packages("usethis") 
```

Next we load R packages required for this script to work. 

```{r load-packages}
# Load some packages
library(available)  # for available()
library(devtools)   #  
library(fs)         # for dir_create()
library(here)       # for here(), i_am()
library(usethis)    # 
```

# Check Candidate Package Names
There are some conventions for naming R packages that you should probably comply 
with. You can use the available package's `available()` function to check 
candidate names for you package. This particular function should actually be
used before you create a new repository from the `rcrpkg.template` because you
want to give the new repository a sensible name from the beginning rather than
have to change it later!

```{r check-pkg-names, eval = FALSE}
# Check availability of the intended package name. 
available("insert.your.package.name")
```

# Create a New Repository Using the Template
On GitHub, you can create a new repository by copying from a template, such as 
my `rcrpkg.template` repository. I think other users should be able to fork and 
clone that repository to their own GitHub accounts, then use it as a template 
themselves. 

- Login to GitHub.
- For users other than myself, I think you would have to start by forking and 
  cloning my `rcrpkg.template` repository to make it available in your own 
  GitHub account. I haven't figured out a way to make the GitHub web interface
  let me try applying a template from another user's account. I'll have to test
  with a colleague to see whether that is possible. 
- Start the process of creating a new repository by visiting 
  https://github.com/new. 
- Use the *Repository template* control to select the `rcrpkg.template`.
- Fill in your new package's name in the *Repository name* box. 
- Add a description.
- Select whether you want the repository to be public or private. 
- Click the *Create repository* button. 

That will add the new repository to your GitHub account.

# Clone the New Repository to Your Computer
Use Git to clone your new repository to your local computer. I use R Studio to
do that with the process described at
https://happygitwithr.com/new-github-first.html#new-rstudio-project-via-git. 

# Rename the R Studio Project File to Match your Repository Name
Your new repository will include an R Studio project file called 
`rcrpkg.template.Rproj` that it inherited from the `rcrpkg.template` repository. 
You will need to rename that file so it matches the name you chose for your new 
repository. So if your new repository is called `MyStudy` you would rename 
`rcrpkg.template.Rproj` to `MyStudy.Rproj`. 

Afterward, use the project file to open your project in R Studio. The remainder 
of this script assumes you have that project file open already. 

# Create a DESCRIPTION File
An R package must have a file called `DESCRIPTION` in the root folder. This file 
holds meta-data about the package. The `create-DESCRIPTION` chunk below will 
create a bare-bones file for you, but then you need to edit that file. In 
particular, you should edit the author, title, and description fields. There is
section of the this script below that will help with the license field. You 
should track the `DESCRIPTION` file in Git. 

```{r create-DESCRIPTION, eval = FALSE}
# Create a DESCRIPTION file. Once it is created, you need to go edit it and 
# then commit it into Git. Make sure you list authors and their roles.
use_description(check_name = TRUE)
```

# Create README.Rmd and README.md Files 
It is very useful for an R pacakge to have README files, which are described 
at https://r-pkgs.org/release.html#important-files. You can use the 
`create-README` chunk below to set up the initial `README.Rmd` file. You can 
edit that file and then knit it to produce the `README.md` file. You should 
track both the `README.Rmd` and `README.md` files in Git. For an example of 
what I put in these files for a pacakge used as a research compendium, see 
https://github.com/sjpierce/SSACHR. 

```{r create-README, eval = FALSE}
# Create a README.Rmd file that will knit to README.md.  
use_readme_rmd()
```

# Create a NEWS.md File 
You should consider using a `NEWS.md` file to record key changes to your 
package. See discussion of such files at https://r-pkgs.org/release.html#news. 
The `create-NEWS` chunk below will set up basic file for you, then you just need 
to edit it as you create new versions of your package. You should track the 
`NEWS.md` file in Git. 

```{r create-NEWS, eval = FALSE}
# Create a NEWS.md file we can use to log changes associated with each version.
# Once it is created, you need to go edit it and then commit it into Git. 
use_news_md()
```

# Create Package-Level Help Documentation
You probably want to do this after you have edited the `DESCRIPTION` file.

```{r create-pkg-help, eval = FALSE}
# Create a dummy .R file in the "R" subfolder that will prompt roxygen to 
# generate basic package-level documentation.
use_package_doc()
```

# Create or Update Help Files and NAMESPACE File

```{r create-NAMESPACE, eval = FALSE}
# Create a NAMESPACE file. Once it is created, you commit it into Git. 
# When DESCRIPTION changes, or when you add new custom functions in the "R"
# subfolder, you should re-run this document() command and commit the updated
# NAMESPACE and help files in the "man" subfolder file to Git. 
document()
```

# Create a Folder for Data
I use the conventional R package structure that calls for storing .Rdata and 
.Rds files in a subfolder called `data`. The `create-data-folder` chunk below
will create the folder for you. 

```{r create-data-placeholder}
# Create the folder
dir_create(path = here("data"))

# Write out the placeholder file. 
Placeholder.txt <- file(here("data/Placeholder.txt"))
Content <- paste("Tracking this file in Git automates creating the folder it", 
                 "is in when the repository is cloned.")
writeLines(text = Content, con = Placeholder.txt)
close(Placeholder.txt)
```

# Create a Folder for External Data Files
I use the conventional R package structure that calls for placing external data
files (i.e., those not stored as .Rdata or .Rds files) in a subfolder called 
`inst/extdata`. The `create-extdata-placeholder` chunk below automates adding a
simple text file called `inst/extdata/Placeholder.txt`. If you want to make sure
someone else cloning your repository will have that folder, just run the chunk 
and then track that file with Git. 

```{r create-extdata-placeholder}
# Create the folder
dir_create(path = here("inst/extdata"))

# Write out the placeholder file. 
Placeholder.txt <- file(here("inst/extdata/Placeholder.txt"))
Content <- paste("Tracking this file in Git automates creating the folder it", 
                 "is in when the repository is cloned.")
writeLines(text = Content, con = Placeholder.txt)
close(Placeholder.txt)
```

# Choose a Software License for the Package
If you do not want to make your package open source (at this time), you can use
the `choose-proprietary-license` chunk below to indicate that you are reserving 
all rights to the copyright holders. You can always change the license later. 

```{r choose-proprietary-license, eval = FALSE}
# Until otherwise decided, we will assume a proprietary license rather than an
# open-source license. 
use_proprietary_license(copyright_holder)
```

On the other hand, if you are using the package as a research compendium that 
you intend to publish, you may want to use the open-source GNU General Public 
License version 3 ([GPL3](https://www.gnu.org/licenses/gpl-3.0.en.html)) or 
or higher for the source code and license all other content under the Creative 
Commons Attribution-ShareAlike 4.0 International 
[CC BY-SA-4.0](https://creativecommons.org/licenses/by-sa/4.0/)) license. I use
those options to promote reproducible research. You can use the 
`choose-GPL3-and-CC-BY-SA-4` chunk below to do that. After running the chunk, 
edit your `DESCRIPTION` file to make the license field read as 
follows: `License: GPL (>= 3) + file LICENSE`. 

```{r choose-GPL3-and-CC-BY-SA-4, eval = FALSE}
# Choose the GPL (>= 3) license
use_gpl_license(version = 3, include_future = TRUE)

# Download the GPL3 license to a file called LICENSE.md
download.file(url = "https://www.gnu.org/licenses/gpl-3.0.txt", 
              destfile = here("LICENSE.md"))

# Write out a LICENSE.note file. 
LICENSE.note <- file(here("LICENSE.note"))
writeLines(c("Source code in this package falls under the GPL (>= 3) license ",
             "(see https://www.gnu.org/licenses/gpl-3.0.en.html or the ",
             "LICENSE.md file). All other contents are licensed under the CC ",
             "BY-SA-4.0 license (see https://creativecommons.org/licenses/by-sa/4.0/ ",
             "or the LICENSE file)."), 
           LICENSE.note)
close(LICENSE.note)

# Write out a LICENSE file. 
download.file(url = "https://creativecommons.org/licenses/by-sa/4.0/legalcode.txt", 
              destfile = here("LICENSE"))
```

# Software Information
Show information about the current software environment. 

``` {r show-version, eval=FALSE}
# Get R and R package version numbers in use.
session_info()
```
